Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta

### 2024-07-30 response/response.tex

\subsection{Comments from Reviewer 1}

\begin{itemize}

\item Reviewer 1 wrote:
\begin{quote}
``No Verification \& Validation of the simulation model is described. At least some discussion regarding the validity of the model has to be included in the case study.''
\end{quote}

\item Reviewer 1 wrote:
\begin{quote}
``No measures of variance are given for the simulation results in Tables 1 and 4. Especially since repeated simulation results were conducted, measures of variance should be included to investigate the statistical significance of the results.''
\end{quote}
Thank you for the comment, standard deviations of mean response times are now given in Table 1 and 4. This is a particularly useful inclusion as the large variance in the response times explains the discrepencies in the simulated and expected survivals.

@@

\item Reviewer 1 wrote:
\begin{quote}
``Algorithm 1: Updating the mutation rate m by subtracting the maximum of m0 and a second number from m0 will lead to a negative (or zero) mutation rate, which does not make sense. Probably should be min instead of max.''
\end{quote}
Thank you, this has now been fixed.

\item Reviewer 1 wrote:
\begin{quote}
``Unclear, why the subchapter 4.1 on survival functions is necessary, since traditional coverage targets are used in the case study and the model formulation does not depend on characteristics of the survival functions. While survival functions are an important performance metric in the reviewers opinion, it is not clear why the reader needs this information to follow the remainder of the paper.''
\end{quote}
We disagree with the reviewer here. In the case study, traditional targets such a mean response times are reported alongside survival measures, based on the survival functions. However, a key part of the paper is the optimisation process, and all scenarios compared are `optimised' scenarios that result from this process. The objective function of the optimisation process maximises survival, based on survival functions. This is evnidenced in Equations 6 and 7, which are component parts of the objective function given in Equation 5. Therefore the survival functions are central to the work presented.

\item Reviewer 1 wrote:
\begin{quote}
``p. 8/9: A short explanation, of what the MINPACK hybrd and hybrj algorithms do and how they work should be included either directly or in the appendix.''
\end{quote}
We have now added this.

\item Reviewer 1 wrote:
\begin{quote}
``There is no discussion of how well the proposed model estimates survival probabilities. This could be easily done by calculating the survival probabilites based on the simulated response times and comparing the simulated survival probabilities with the ones calculated by the MESLMHPHF. It is also not clear, why e.g. in Table 4 the expected Survival from the MESLMHPHF was used instead of the more realistic survival based on the simulation results.''
\end{quote}
We have now added new subsection comparing survival obtained through the MESLMHPHF and through the simulation. There are discrepancies between the two, and in the new subsection we discuss reasons for this, and offer some directions of future work that could help in alleviating some of these problems.

@@

\item Reviewer 2 wrote:
\begin{quote}
``Is the survival function the same for $K_a$ patients regardless of which vehicle visits them? What is the rationale for this? It seems that primary vehicles would be better prepared to deal with complications, etc.''
\end{quote}
This is an interesting suggestion. Currently we model the survival function to be independent of the vehicle type, that is $s_k$ is the same regardless if the vehicle that arrived is a primary or secondary vehicle. As you note, this may not be realistic, and we may want to be able to model situations where there are different survival functions for each vehicle type. However, this becomes more complicated when both vehicles arrive, but at different times - there will be an interaction between the functions that requires more investigation and is beyond the scope of this study. These scenarios are interesting, and we have added a sentence in the paper explaining this as further work.

\item Reviewer 2 wrote:
\begin{quote}
``Have you given any consideration to using an exact method to solve the optimisation problem solved in Algorithm 1? Is there any alternative in the literature?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P11 L43 – What is the rationale of using a triangular distribution for travel times? In general, what is the basis of all the travel time assumptions presented in this section?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P12 5.4 – For patients that might not need to travel to a hospital, would they still need a primary vehicle after a secondary vehicle tends to them? If not the logic would not be separate entirely.''
\end{quote}
This is a very good point, and is similar to the situation when a primary vehicle reaches the patient before the secondary vehicle. Ideally in these cases the later vehicle would turn around and be free to pick up another patient. However one assumption of the model is that en-route communications are difficult and do not occur. We have now emphasised this in the text.

\item Reviewer 2 wrote:
\begin{quote}
``It seems there are two conflicting methods here, the optmisiation with the EA, which uses some averages, etc. And a more comprehensive DES simulation; however the latter is not used for the optimisation. Did you consider simulation optimisation or simheuristics to address this?''
\end{quote}
We did not consider using the simulation as part of the optimisation process, and is beyong the scope of the paper. However, we think this is a great idea, and have discussed this as future work in the paper, as a tool to give greater consideration to the variability in the system.

### 2024-07-30 tex/main.tex
address comments on using simulation survival, adding variability measures

\subsection{Survival Functions}\label{sec:survival}
A key concept here is the survival curve of a patient.
In reality, some emergency incidents do not result in a substantive
deterioration of a patient’s status over time; however, in all situations,
there is a reasonable cut-off beyond which a patient should not expect to have
to wait for care, and mixture of theoretical survival functions and step
functions can be utilised.
Survival probabilities for critical incidents, calculated from a theoretical
monotonically decaying survival function reported in the literature
\citep{Valenzuela20001206} are used to demonstrate an attainable level of
success from a response. One particular survival curve $s(t)$ of
Equation~\ref{eqn:survival_curve}, represents survival until hospital
discharge following cardiac arrest; its origins are explained in detail by
\citet{Knight2012918}, and gives the probability of survival if seen within a
time $t$ in the form of a logistic function. Figure~\ref{fig:survivalfunction}
shows the difference between using this survival curve and a hard cut-off of 8
minutes.
However, hard cut-off curves like Equation~\ref{eqn:survival_cutoff}, with a
cutoff of $L$ can still be used to represent meeting artificially selected
targets, for example for transportation jobs.

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{img/Survival_Function.pdf}
\caption{Survival function $s(t)$, estimated by \citet{Valenzuela20001206}
compared with $s_L(t)$ with a hard cutoff of $L=8$, represented as a step
function.}
\label{fig:survivalfunction}
\end{figure}

@@

Exact synchronicity considerations are shown in Figure~\ref{fig:sequential_logic}:
\begin{itemize}
  \item whenever the secondary vehicle reaches the patient before the primary
  has left, vehicle, they remain with the patient until the primary vehicle
  leaves;
  \item if secondary vehicle arrives after the primary vehicle has left, they
  will immediately return to their stations as they are not needed at the scene;
  \item if the expected time for the secondary vehicle to reach the patient
  exceeds the expected time for the primary vehicle to reach the patient then
  the secondary vehicle is not deployed, and so that transit job would be
  abandoned in the secondary simulation (although still seen by a primary
  vehicle in the initial simulation). We assume that secondary vehicles must
  still reach the pickup locations, due to potential difficulties in
  communication en-route.
\end{itemize}

@@

\begin{table}
\begin{tabular}{rrrr}
\toprule
Allocation & Baseline & Grid 3km & Grid 5km \\
\midrule
Number of EAs & 81 & 70 & 71 \\
Number of RRVs & 13 & 0 & 0 \\
Ambulance Utilisation & 30.99\% & 36.58\% & 35.79\% \\
RRV Utilisation & 22.47\% & - & - \\
Mean Response Time (mins) & 17.69 & 22.29 & 24.05 \\
MESLMHPHF Objective & 98.34\% & 87.95\% & 85.64\% \\
Expected A1 Survival & 26.05\% & 13.64\% & 12.23\% \\
Percent Abandoned & 0.00\% & 0.52\% & 0.93\% \\
\bottomrule
\end{tabular}
\caption{Calculated KPIs for the current and proposed grid allocations.}
\label{tbl:current_grid_results}
\end{table}

@@

Table~\ref{tbl:current_optimal_compare} compares the MESLMHPHF objective
function value and simulation derived KPIs between the current and the optimised
allocations, for each demand scenario. It can be seen that, as expected, as
demand increases then utilisation of primary and secondary vehicles increase, as
the ambulance service is busier. Similarly, mean response time increases with
demand, this is due to the vehicles being busier, and therefore less chance that
the most optimally placed vehicle is dispatched. This is confirmed by the
percentage of calls abandoned increasing with demand, showing ambulance
unavailability.
Looking at the expected survival, that is the value of the MESLMHPHF objective
function, we see that using an better allocation does increase the survival,
and that the drop in survival due to demand increases is not as severe when
using an improved allocation in comparison to the current allocation. We can
also see that in particular A1 patients are benefiting from the improved
allocations, as these are the patients most sensitive to small fluctuations in
response time, due to their survival function Equation~\ref{eqn:survival_curve}.
It is worth noting that for the simulation based KPIs (mean response times and
percent abandoned), there is not too much
difference between the currently used allocation and the better allocation,
showing that these KPIs may not be good approximations of survival.

@@

>>>>

\subsection{Survival Functions}\label{sec:survival}
A key concept here is the survival curve of a patient.
In reality, some emergency incidents do not result in a substantive
deterioration of a patient’s status over time; however, in all situations,
there is a reasonable cut-off beyond which a patient should not expect to have
to wait for care, and mixture of theoretical survival functions and step
functions can be utilised.
Survival probabilities for critical incidents, calculated from a theoretical
monotonically decaying survival function reported in the literature
\citep{Valenzuela20001206} are used to demonstrate an attainable level of
success from a response. One particular survival curve $s(t)$ of
Equation~\ref{eqn:survival_curve}, represents survival until hospital
discharge following cardiac arrest; its origins are explained in detail by
\citet{Knight2012918}, and gives the probability of survival if seen within a
time $t$ in the form of a logistic function. Figure~\ref{fig:survivalfunction}
shows the difference between using this survival curve and a hard cut-off of 8
minutes.
However, hard cut-off curves like Equation~\ref{eqn:survival_cutoff}, with a
cut-off of $L$ can still be used to represent meeting artificially selected
targets, for example for transportation jobs.

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{img/Survival_Function.pdf}
\caption{Survival function $s(t)$, estimated by \citet{Valenzuela20001206}
compared with $s_L(t)$ with a hard cut-off of $L=8$, represented as a step
function.}
\label{fig:survivalfunction}
\end{figure}

@@

Exact synchronicity considerations are shown in Figure~\ref{fig:sequential_logic}:
\begin{itemize}
  \item whenever the secondary vehicle reaches the patient before the primary
  has left, vehicle, they remain with the patient until the primary vehicle
  leaves;
  \item if secondary vehicle arrives after the primary vehicle has left, they
  will immediately return to their stations as they are not needed at the scene;
  \item if the expected time for the secondary vehicle to reach the patient
  exceeds the expected time for the primary vehicle to reach the patient then
  the secondary vehicle is not deployed, and so that transit job would be
  abandoned in the secondary simulation (although still seen by a primary
  vehicle in the initial simulation). We assume that secondary vehicles must
  still reach the pickup locations, due to potential difficulties in
  communication en-route. \textcolor{red}{Consider also the situation when a
  secondary vehicle arrives and the patient does not require a hospital, with
  better en-route communication this would mean the primary vehicle turns
  around, however without en-route communication which is assumed here, the
  primary vehicle must still reach the patient pick-up location.}
\end{itemize}

@@

\begin{table}
\begin{tabular}{rrrr}
\toprule
Allocation & Baseline & Grid 3km & Grid 5km \\
\midrule
Number of EAs & 81 & 70 & 71 \\
Number of RRVs & 13 & 0 & 0 \\
Ambulance Utilisation & 30.99\% & 36.58\% & 35.79\% \\
RRV Utilisation & 22.47\% & - & - \\
Mean Response Time (mins) & 17.69 & 22.29 & 24.05 \\
\textcolor{red}{StDev Response Time (mins)} & \textcolor{red}{6.98} & \textcolor{red}{8.02} & \textcolor{red}{7.94} \\
MESLMHPHF Objective & 98.34\% & 87.95\% & 85.64\% \\
Expected A1 Survival & 26.05\% & 13.64\% & 12.23\% \\
Percent Abandoned & 0.00\% & 0.52\% & 0.93\% \\
\bottomrule
\end{tabular}
\caption{Calculated KPIs for the current and proposed grid allocations.}
\label{tbl:current_grid_results}
\end{table}

@@

Table~\ref{tbl:current_optimal_compare} compares the MESLMHPHF objective
function value and simulation derived KPIs between the current and the optimised
allocations, for each demand scenario. It can be seen that, as expected, as
demand increases then utilisation of primary and secondary vehicles increase, as
the ambulance service is busier. Similarly, mean response time increases with
demand, this is due to the vehicles being busier, and therefore less chance that
the most optimally placed vehicle is dispatched. This is confirmed by the
percentage of calls abandoned increasing with demand, showing ambulance
unavailability.
Looking at the expected survival, that is the value of the MESLMHPHF objective
function, we see that using an better allocation does increase the survival,
and that the drop in survival due to demand increases is not as severe when
using an improved allocation in comparison to the current allocation. We can
also see that in particular A1 patients are benefiting from the improved
allocations, as these are the patients most sensitive to small fluctuations in
response time, due to their survival function Equation~\ref{eqn:survival_curve}.
\textcolor{red}{There is a large discrepancy between the expected survival
measures calculated from the MESLMHPHF objective, and the observed survival from
the simulation. This is discussed further in the next section.}
It is worth noting that for the simulation based KPIs (mean response times and
percent abandoned), there is not too much
difference between the currently used allocation and the better allocation,
showing that these KPIs may not be good approximations of survival.

@@

\textcolor{red}{
\subsection{Comparing Expected \& Simulated Survival Measures}
In Table~\ref{tbl:current_optimal_compare} it can be seen that there is a large
discrepancy between the expected survival, both overall and of A1 patients,
computed using the MESLMHPHF described in Section~\ref{sec:meslmhphf} and
calculated using the simulation. This is a consequence of the variability in
response times, which is explicitly modelled by the simulation but not captured
by the MESLMHPHF. As seen in the Table, the standard deviations of the response
times are relatively large in comparison to the mean response times, and so this
effect would be substantial here.}

\textcolor{red}{
The discrepancies are exacerbated by the use of cut-off survival function curves
for most of the patients: consider a survival function with cut-off $L$, with
response time below $L$ the patient's chance of survival is 0, and above $L$ the
patient's chance of survival is guaranteed. The optimisation will find a
solution such that will ensure that the expected response time is above $L$,
possibly as low as $L + \epsilon$, but will not attempt to push the expected
response time above this, as it will have no effect on the expected survival.
However considering variability of mean response times, a considerable
proportion of the response times will lie below the expected value, and so
possibly below $L$, resulting in worse survival probabilities, as observed in
Table~\ref{tbl:current_optimal_compare}. Therefore the distributions chosen to
model travel times can have a large impact on the survival measures, in this
case the naive Triangular distribution.}

\textcolor{red}{
This leads to a number of avenues of future work to improve this and overcome
this effect. Variability could be explicitly modelled within the MESLMHPHF, such
that it is not only maximising \textit{expected} survival, but considers some
percentiles of guaranteed survival. Another avenue would be to use the
simulation explicitly in the optimisation process, incorporating the simulation
into the objective function and using simulation-optimisation techniques to
improve runtime efficiency. However, as discussed in Section~\ref{sec:survival},
maximising expected survival of all patients can lead to a fairer allocation of
resources, with difficult to reach patients being given equal consideration to
closer, easier to reach patients. When considering the variability of travel
times in the optimisation process, difficult to reach patients may be ignored, indistinguishable from patients who happen to sample a response time at the
extreme tail of a given distribution curve.}

\textcolor{red}{
Therefore, due to this interpretation of fairness of allocations, and due to the
naive choice of travel time distributions, including difficulties in being able
to validate such a model and the choices made, we consider only the MESLMHPHF
objective from now on.
}
